Individual and Differential Harm in Redistricting

### 2022-05-09 paper/paper.Rmd
redistricting -> districting

-   **we argue that individual standing/harm is an excellent lens to unify the study and measurement of redistricting plans**

-   core idea: counterfactuals + voter choice.
    Is a voter more likely to be represented by the candidate of their choice under a different redistricting plan?
    If so, they've been harmed.

@@

For example, the algorithms of @carter2019 and @mccartan2020 sample from the maximum-entropy distribution of districting plans which are geographically contiguous, have district populations within a deviation from equality, and which have a certain average compactness value (according to a particular graph-theoretic compactness measure).
These maximum-entropy distributions put as few conditions as possible on the districting sampling process, as long as a set of constraints imposed by the researcher are met.
If these constraints are precisely those which are required by law or traditional redistricting criteria, then samples from this distribution provide a natural set of counterfactual districting plans, uninfluenced by superfluous partisan considerations, to use in estimating $H(q)$ and $\Delta H(q)$.
In practice, the translation from tradition and law into mathematical formulations of constraints is not always clear or even well-defined; there is certainly an element of subjectivity in sampling a set of redistricting plans.
Nonetheless, sampling methods are widely used in practice.

@@

But as those authors recognize, in evaluating a districting plan, it is critical to consider "if there is a fairer alternative." The inability of partisan bias and other existing measures to account for the space of counterfactual redistricting plans means that none of them may be interpreted in a vacuum.
No particular value can be used as a threshold past which a plan becomes a gerrymander, because the measures confound the map-drawer's bias with the bias inherent to the district-based electoral system.

>>>>

-   **we argue that individual standing/harm is an excellent lens to unify the study and measurement of districting plans**

-   core idea: counterfactuals + voter choice.
    Is a voter more likely to be represented by the candidate of their choice under a different districting plan?
    If so, they've been harmed.

@@

For example, the algorithms of @carter2019 and @mccartan2020 sample from the maximum-entropy distribution of districting plans which are geographically contiguous, have district populations within a deviation from equality, and which have a certain average compactness value (according to a particular graph-theoretic compactness measure).
These maximum-entropy distributions put as few conditions as possible on the districting sampling process, as long as a set of constraints imposed by the researcher are met.
If these constraints are precisely those which are required by law or traditional redistricting criteria, then samples from this distribution provide a natural set of counterfactual districting plans, uninfluenced by superfluous partisan considerations, to use in estimating $H(q)$ and $\Delta H(q)$.
In practice, the translation from tradition and law into mathematical formulations of constraints is not always clear or even well-defined; there is certainly an element of subjectivity in sampling a set of districting plans.
Nonetheless, sampling methods are widely used in practice.

@@

But as those authors recognize, in evaluating a districting plan, it is critical to consider "if there is a fairer alternative." The inability of partisan bias and other existing measures to account for the space of counterfactual districting plans means that none of them may be interpreted in a vacuum.
No particular value can be used as a threshold past which a plan becomes a gerrymander, because the measures confound the map-drawer's bias with the bias inherent to the district-based electoral system.
